---
title: "Toronto Island Traffic Analysis"
subtitle: "."
author: 
  -Bo Tang
thanks: "Code and data are available at: [RayanAlim/Toronto_flooding_analysis](https://github.com/RayanAlim/Toronto_flooding_analysis)"
date: today
date-format: long
abstract: ""

format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(palmerpenguins)
monthly_sales <- read.csv("D:/Documents/starter_folder-main/data/analysis_data/analysis_data_month.csv")
```

# Introduction

The article looks at the number of Toronto Island Ferry Tickets sold in the last year to figure out when the main time period for the public to go to the island is, i.e., the peak time.

# Data {#sec-data}



```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(monthly_sales, aes(x = month, y = total_sales)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Monthly Sales Comparison", x = "month", y = "monthly sales") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Results {#sec-results}


## Voter Turnout


## Poll Results


## Response Rate


# Discussion {#sec-dis}

## Key Findings


## Limitations and Future Work

\newpage

\appendix

# Appendix {-}


\newpage


# References